---
title: "xportrjson"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{xportrjson}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Objectives

1. Read in a dataset-JSON DM file
2. Add a new variable to the dataset
3. Update metadata within JSON file
4. Write out the dataset-JSON file


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

library(DT)
```

## Read in a dataset-JSON DM file

Below we use the `read_dataset_json` to read into our R session a Demographics dataset-JSON file via a url. 


```{r setup, message = FALSE}
library(dplyr)
library(jsonlite)
library(xportr)
library(xportrjson)

dm <- read_dataset_json(url("https://raw.githubusercontent.com/lexjansen/sas-papers/master/pharmasug-2022/json/sdtm/dm.json"))
```

```{r, echo=FALSE}
glimpse(dm)
```

The `read_dataset_json` also reads in the metadata from the JSON file and applies it to the dataframe.  For example, the variable `RFXSTDTC` has the label `Date/Time of First Study Treatment` applied to it.
 
```{r}
attributes(dm$RFXSTDTC)
```

To get a better idea of what is in the JSON metadata file for the DM data, lets take a quick look.

```{r}
raw_dm <- jsonlite::fromJSON(url("https://raw.githubusercontent.com/lexjansen/sas-papers/master/pharmasug-2022/json/sdtm/dm.json"), simplifyVector = TRUE)

head(raw_dm$clinicalData$itemGroupData[[1]]$items, 5)
```

## Update metadata within JSON file

The JSON file has the metadata associated with the JSON dataset.  If we would like to add variables to the JSON file while using R we will need to do three tasks.

1) Pull out current metadata for DM dataset
2) Add new variable metadata

### Pull out current metadata for DM dataset

```{r}
metadata_current <- raw_dm$clinicalData$itemGroupData[[1]]$items

head(metadata_current, 5)
```

### Add new variable metadata

```{r}
metadata_updated <- data.frame(
  name = c("TRT01P", "TRT01A"),
  OID = c("IT.ADSL.TRT01P", "IT.ADSL.TRT01A"),
  length = c(28, 28),
  type = c("string", "string"),
  label = c("Planned Treatment for Period 01", "Actual Treatment for Period 01")
) %>% 
  bind_rows(metadata_current) %>% 
  rename(variable=name)
```


## Add a new variable to the dataset

Now we would like to add a new variable and update the metadata associated with this variable on the R dataframe.  We will use the `{xportr}` functions to apply the length and label to the R dataframe.

```{r, message = FALSE, warning = FALSE}
adsl <- dm %>%
  mutate(TRT01P = ARM, TRT01A = ACTARM) %>% 
  xportr_length(metadata_updated) %>% 
  xportr_label(metadata_updated)

attributes(adsl$TRT01P)
```

## Write out an updated JSON (To finish)

TODO: Figure out how to put together updated metadata and data into one JSON file
TODO: Write out and archive somewhere accessible.

```{r}
write_item_data(adsl)
```
